---
title: "Data Manipulation"
output:
  html_document:
      toc: true
      toc_float: true
---

```{r setup, include=FALSE}

knitr::opts_chunk$set(echo = TRUE)

library(tidyverse)

```

### Basic Data Frames

A `tibble`, or `tbl_df`, is a modern reimagining of the data.frame.
Let's start with a simple data frame.

```{r tibble}

# Anytime appling a random, set a seed for reproductivity
set.seed(20191122)

# Let's create a data frame including:
# 1. A random sample of size 10 from a uniform [0, 5] distribution
# 2. A logical vector indicating whether elements of the sample are greater than 2
# 3. A (length-10) character vector
# 4. A (length-10) factor vector
data1 = tibble(
  vec_unif = runif(10, min = 0, max = 5),
  vec_logical = vec_unif > 2,
  vec_char = c("My", "name", "is", "Sha", "Tao", "and", "I", "like", "Data", "Science"),
  vec_factor = factor(c("low", "low", "low", "high", "high", "low", "low", "low", "high", "high"))
)
data1

# Try to take the mean of the numeric variable
mean(data1$vec_unif)

# Try to take the mean of the logical variable
mean(data1$vec_logical)

```

The first two vectors work because a uniform [0, 5] distribution is numeric, and "true" and "false" in logical vectors are treated as 1 and 0.
Means cannot be taken from the character vector and factor vector.

### Piping

` %>% ` is one of the most important function in R comparing to SAS, and it's one of the main reasons that I prefer to use R.

### Common Functions

We'll use some common functions in data manipulation.
First import dataset we used last time:

```{r import}

litters_data = read_csv(file = "./data/r_data_import/FAS_litters.csv") %>% 
  janitor::clean_names()

pulse_data = haven::read_sas("./data/r_data_import/public_pulse_data.sas7bdat") %>%
  janitor::clean_names()

```


#### `select`

We can select desired columns by naming them.

```{r select1}

litters_data %>% 
  select(group, litter_number, gd0_weight, pups_born_alive) %>% 
  head()

```

We can select a group of desired columns.

```{r select2}

litters_data %>%
  select(group:gd_of_birth) %>% 
  head()

```

We can select desired columns by removing unwanted ones.

```{r select3}

litters_data %>% 
  select(-pups_survive) %>% 
  head()

```

We can rename variables as part of the select process.

```{r select4}

litters_data %>% 
  select(group1 = group, litter_number2 = litter_number) %>% 
  head()

```

If all we want to do is renaming, then we can use `rename`.

```{r select5}

litters_data %>% 
  rename(group1 = group, litter_number2 = litter_number)

```

Another common function that is quite useful is `everything()`, we can use it to rearrange the column order easily.

```{r select6}

litters_data %>% 
  select(litter_number, group, everything())

```

Check the functions `starts_with()`, `ends_with()`, and `contains()` which are handy in large datasets.
In addition,  `select` will export a dataframe even if you only select one column. To pull a single variable, use function `pull`.

#### `filter`

We will often filter using comparison operators (`>`, `>=`, `<`, `<=`, `==`, and `!=`). We may also use `%in%` to detect if values appear in a set, and `is.na()` to find missing values. 
The results of comparisons are logical – the statement is `TRUE` or `FALSE`, and can be combined with other comparisons using the logical operators `&` and `|`, or negated using `!`.

Here are some of the filters we may use in the example dataset:

* `gd_of_birth == 20`
* `pups_born_alive >= 2`
* `pups_survive != 4`
* `!(pups_survive == 4)`
* `group %in% c("Con7", "Con8")`
* `group == "Con7" & gd_of_birth == 20`

`drop_na()` is a good function to omit missing observations.

* `drop_na(litters_data)` will remove any row with a missing value.
* `drop_na(litters_data, wt_increase)` will remove rows for which wt_increase is missing.

#### `mutate`

`mutate` is the function we use to change or create variables. Here's an example.

```{r mutate1}

litters_data %>%
  mutate(wt_gain = gd18_weight - gd0_weight,
  group = str_to_lower(group)
  ) %>% 
  head()

```

`stringr` is a great package. `str_to_lower()` convert variables to all lower case, there are also functions `str_to_upper()`, `str_to_title()`, `str_to_sentence()`.

#### `arrange`

We can arrange the rows in your data according to the values in one or more columns.

```{r arrange}

litters_data %>% 
  arrange(group, pups_born_alive) %>% 
  head(10)

```

#### Transposing wide to long, `gather` & `pivot_longer`

To make dataset "tidy", we need to do some transposing. I used to use `gather` function a lot, but nowadays it seems people use `pivot_longer` more often. I'll introduce both of them here.

```{r transpose1}

pulse_tidy1 = 
  pulse_data %>% 
  gather(key = visit, value = bdi, bdi_score_bl:bdi_score_12m) %>% 
  arrange(id)

pulse_tidy1 %>% 
  head(10)

```

As you can see that the wide "untidy" form becomes a long "tidy" form. For the variable `visit`, the original column names were informative but we probably don’t need to keep the `bdi_score_ prefix` in each case.

```{r transpose2}

pulse_tidy1 = 
  pulse_tidy1 %>% 
  mutate(visit = str_replace(visit, "bdi_score_", "")) %>% 
  arrange(id)

pulse_tidy1 %>% 
  head(10)

```

Now the data looks much better. However, it seems that `sex` and `visit` are in the form of character but not factor. Also, we may want to change `bl` to `00m` for consistency across visits. 

```{r transpose3}

pulse_tidy1 = 
  pulse_tidy1 %>% 
  mutate(sex = factor(sex),
         visit = replace(visit, visit == "bl", "00m"),
         visit = factor(visit, levels = str_c(c("00", "01", "06", "12"), "m"))) %>%
  arrange(id, visit)

pulse_tidy1 %>% 
  head(10)

```

Now the dataset looks even better, and is ready to be analyzed. Let's put the steps together:

```{r transpose4}

pulse_tidy1 = 
  pulse_data %>% 
  gather(key = visit, value = bdi, bdi_score_bl:bdi_score_12m) %>% 
  mutate(sex = factor(sex),
         visit = str_replace(visit, "bdi_score_", ""),
         visit = replace(visit, visit == "bl", "00m"),
         visit = factor(visit, levels = str_c(c("00", "01", "06", "12"), "m"))) %>% 
  arrange(id, visit)

pulse_tidy1 %>% 
  head(10)

```

What about `pivot_longer`?

```{r transpose5}

pulse_tidy2 = 
  pulse_data %>%
  pivot_longer(bdi_score_bl:bdi_score_12m,
               names_to = "visit",
               names_prefix = "bdi_score_",
               values_to = "bdi") %>%
  select(id, visit, everything()) %>%
  mutate(
    visit = replace(visit, visit == "bl", "00m"),
    visit = factor(visit, levels = str_c(c("00", "01", "06", "12"), "m"))) %>%
  arrange(id, visit)

```
















