---
title: "Data Manipulation"
output:
  html_document:
      toc: true
      toc_float: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r problem1_1 }

set.seed(1)

# Create a data frame including:
# 1. A random sample of size 10 from a uniform [0, 5] distribution
# 2. A logical vector indicating whether elements of the sample are greater than 2
# 3. A (length-10) character vector
# 4. A (length-10) factor vector
data1 = tibble(
  vec_unif = runif(10, min = 0, max = 5),
  vec_logical = vec_unif > 2,
  vec_char = c("My", "name", "is", "Sha", "Tao", "and", "I", "like", "Data", "Science"),
  vec_factor = factor(c("low", "low", "low", "high", "high", "low", "low", "low", "high", "high"))
)

# Try to take the mean of the numeric variable
mean(data1$vec_unif)

# Try to take the mean of the logical variable
mean(data1$vec_logical)

# mean(data1$vec_char) doesn't work
# mean(vec_factor) doesn't work
```

The first two vectors work because a uniform [0, 5] distribution is numeric, and "true" and "false" in logical vectors are treated as 1 and 0. \
Means cannot be taken from the character vector and factor vector.


## Problem 1_2 The chunk below converts variables from one type to another.

```{r problem 1_2, results = "hide"}

#applies the as.numeric function to the logical, character, and factor variables
as.numeric(data1$vec_logical)
as.numeric(data1$vec_char)
as.numeric(data1$vec_factor)

```

logical vector can be convert to numeric, and "true" and "flase"" are treated as 1 and 0.\
For factor vectors, R re-orders the factors alphabetically, and assigns numeric numbers to each level. In my case, it assigns 1 to "high" and 2 to "low"."\
Function as.numeric cannot treat character vector in my case and it returns 10 "NA"s. However, if I use number "1", "2" as my character vector, then it will return as numeric vector 1, 2.


## Problem 1_3 The chunk below does a two-step converting.

```{r problem 1_3}

# Convert character variable from character to factor to numeric
as.numeric(as.factor(data1$vec_char))

# Convert factor variable from factor to character to numeric
as.numeric(as.character(data1$vec_factor))
```

In the first case, function as.factor converts my character vector to a factor vector. Then function as.numeric reorders the  factors alphabetically, and assigns numeric numbers to each level.\
In the second case, function as.character converts my factor vector to a character vector, thus cannot be treated and returns 10 "NA"s.


## Problem 2_1 The chunks below creates a data frame for later plotting.

```{r problem 2_1}

set.seed(2)

# Create a random sample x of size 1000 from a standard Normal distribution
x = rnorm(1000)

# Create a random sample y of size 1000 from a standard Normal distribution
y = rnorm(1000)

# Create a data frame including:
# 1. x  2. y
# 3. A logical vector indicating whether the x + y > 0
# 4. A numeric vector by coercing the above logical vector
# 5. A factor vector created by coercing the above logical vector
data2 = tibble(
  x,
  y,
  xy_logical = x + y > 0,
  xy_numeric = as.numeric(xy_logical),
  xy_factor = as.factor(xy_numeric)
)

```

The dataset "data2" is a `r ncol(data2)` by `r nrow(data2)` table.\
The mean of the variable x is `r round(mean(data2$x), digit = 2)`, while the median of the variable x is `r round(median(data2$x), digit = 2)`.\
There is `r (sum(data2$xy_numeric)/1000)*100`% of cases which the logical vector "x + y > 0" is true.

